Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-15 main.tex
wip

\begin{itemize}
\item
  Idea of accessibility: set of data memory addresses that can legally be read or written at a particular execution point.
  (Assume instruction memory is distinct.)
\item
  Stack protection intuition in a nutshell: caller's memory data is not accessible to callee, although callee's memory data is accessible to caller.
  This gives both an integrity property (callee cannot affect caller's memory) and a secrecy property
  (caller's memory cannot affect callee).
  Can generalize to other control situations, such as coroutines, threads, compartments,...
\item
  In NI terms, protected programs are oblivious to variations in memory outside of their accessible set.
  Unlike standard NI, the accessible set changes dynamically, so we cannot just frame the property in terms of
  variant start states. [Is this true? Is it a significant difference? Are there other significant differences?]
\end{itemize}

We start with a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
We assume that memory is divided into two disjoint regions, one containing the code, the other the stack.
Code begins at address 0; the stack begins at address $B$ and grows in the direction of increasing addresses.

The part of the stack between $B$ and the stack pointer \textsc{sp} is structured into a sequence of consecutive
\emph{frames}. Each frame consists of a single return address and a single word of local data.

  \begin {itemize}
  \item Memory is divided into an instruction region and a data region. The data region contains a single contiguous stack, divided into frames, each associated with a procedure activation. (The stack grows in the direction of increasing addresses.)
  \item Each frame consists of a saved return address and a single word of local data. These [both??] form the entirety of the procedure's memory data, and are the only thing we need to track when considering accessibility.
  \item Registers other than RA, SP are not affected by calls or returns and hence are the (only) way to pass values between caller and callee.
  \item Instructions serving as calls are statically marked as such in the code. [The entry and exit sequences must also be marked???]
  \item The code for each procedure is bracketed by fixed entry and exit sequences (details to follow), which save/restore the RA and
    adjust the SP. [Are these assumptions or checked facts?]
  \item During the body of the procedure, the single word of local data is accessed via SP, which does not change (except during a call, of course).  The stored RA in the frame never changes. [I'm confused about these. They are checked facts, not assumptions.]
  \end {itemize}
  In this model, the definition of accessible memory is simply all (data) memory at addresses $\geq pc$.
\begin{itemize}
\item
  Describe conjecture 2 property in terms of this model: informally and formally.
\item
  Define abstract state descriptions.
\item
  Define step consistency.
\item
  Discuss relationship between state consistency and conjecture 2 property.  (Former is: presumably stronger; hopefully not
  too strong that it produces lots of undesired test failures; surely easier to test).

>>>>

Here we consider stack protection as an instance of generalized interference \apt{or whatever it's called}.
Our notion of stack protection includes aspects both of control integrity and data security.
In the terminology of Skorstengaard et al.~\cite{Skorstengaard+19} we enforce:

\begin{itemize}
\item \emph{Well-bracketed control flow (WBCF)}, which says that ``invoked functions must either return to their callers,
  invoke other functions themselves, or diverge.''
\item \emph{Local state encapsulation (LSE)}, which says that ``when a function invokes another function, its local variables
  (saved on its stack frame) will not be read or modified until the invoked function returns.''
\end{itemize}
\apt{Should rephrase (and improve!) these in our own words.}

These properties are essentially orthogonal: we could enforce WBCF without LSE, and variants of
LSE could apply to control constructs beyond simple procedure calls (e.g. exceptions, coroutines, threads, monitors, etc.).

In terms of NI....
\apt{Perhaps the NI-like property is just LSE, but it holds only if WBCF holds?}

To make these ideas concrete in this paper, we adopt a very simple (and unrealistic) set of assumptions/conventions for stack frames and procedure calls.
Each procedure stack frame contains just one local variable (plus possible control information), which is pointed to by
the SP whenever the PC is within the procedure body (excluding entry and exit sequences). Our WBCF property says that
when a procedure is called from a given PC at a given SP, it eventually returns (if at all) to the next
address following the PC with the same SP. Our LSE-like property says that addresses between the base of the stack
up to (but not including) the current SP cannot be read or written.
\apt{Not clear why we should bother allowing addresses beyond the current top of stack to be accessible.}
In particular, this implies that upon return from a call,
the value of the local variable will be unchanged; moreover, the values of local variables cannot influence the
behavior of callees. Parameters and return values are passed in registers, which (except for SP and RA) do not change
during call or return.

{\bf Getting more precise about the property.}

Assumptions on code and environment\apt{trying to be as minimal as possible}:
\begin{itemize}
\item 
  Memory is divided into two disjoint regions, one containing the code, the other the stack.
  Code begins at address 0; the stack begins at address B and grows up, in the direction of increasing addresses.
\item
  The part of the stack between $B$ and the stack pointer \textsc{sp} (inclusive) is structured into a contiguous sequence of
  \emph{frames}. Each frame contains a single word of local data, possibly preceded by control information. \apt{TMI?}
  The initial value of \textsc{sp} is $B-1$.
\item
  Code is assumed to be be annotated to indicate (a) instructions representing calls, (b) instructions forming the \emph{entry sequence} of
  a procedure, and (c) instructions forming the \emph{return sequence} of a procedure.
\item
  Execution of the entry sequence pushes a new frame on the stack, adjusting \textsc{sp} to point to that
  frame's data word.  Execution of the return sequence pops the frame on top of the stack by readjusting \textsc{sp}.
  \apt{TMI!?}
  Outside of the entry and return sequences, no instruction alters \textsc{sp}.
\item
  The call instruction and entry and return sequences do not alter any registers except \textsc{sp} and one
  particular register named \textsc{ra}.
\end{itemize}

DO HERE: Formal description of conjecture 2 property.  Or maybe we no longer really care about this particular property if we have something for conjecture 3?

An (abstract) \emph{state description} $D$ consists of
\begin{itemize}
\item
  A stack $\dstk{D}$ of addresses, representing the return addresses of the current procedure and its callers. 
  \apt{No need to remember the sp's, since they can be computed given that we know the fixed frame size.}
  \apt{No need for $\dpcd{D}$ since it is just the size of $\dstk{D}$.}
\item
  For each memory location $a \geq B$, a natural number $\ddep{D}{a}$, representing the depth
  of the frame to which $a$ belongs ($\infty$ if none).
\end{itemize}

The \emph{initial state description} $\dinit$ has $\dstk{\dinit} = \empstack$ and $\ddep{\dinit}{a} = \infty$ for all $a \geq B$.

An address $a$ is \emph{accessible} with respect to a state $D$, written $\access{D}{a}$, if
$a \geq B$ and $\ddep{D}{a} \geq \norm{\dstk{D}}$.

TODO:
\begin{itemize}
\item
  Define step relation.
\item
  Define step consistency.
\item   
  Discuss relationship between state consistency and top-level property.  (Former is: presumably stronger; hopefully not
  too strong that it produces lots of undesired test failures; surely easier to test).
